---
title: "Clean Code in R"
subtitle: "How to write human-readabe code?"
author: "Lenard Dome"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
  xaringan::moon_reader:
    css: ["libs/default.css", "libs/default-fonts.css"]
    seal: true 
    self_contained: false
    nature:
      ratio: "21:11"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: middle left inverse

```{r, include = FALSE}
library(emo)
library(fontawesome)
```

# Today we will focus on:

- `r emo::ji("document")` Documentations and Commenting Code
- `r fa("code")` Readable Code
- `r emo::ji("robot")` Functional Programming and Programming with Functions

---

class: middle left inverse

# Documentations

**Documentations explains what you do and how**

- Manual for using your code and reproducing
- Documenting functions: use [standard R documentation](https://cran.r-project.org/doc/manuals/R-exts.html#Writing-R-documentation-files)
- Documenting your analysis, use [vignettes](https://r-pkgs.org/vignettes.html).
- Examples of Vignettes: [BEST package](https://cran.r-project.org/web/packages/BEST/vignettes/BEST.pdf), [dplyr of the tidyverse family](https://cran.r-project.org/web/packages/dplyr/vignettes/base.html).

---

class: middle left inverse

# Commenting Code

**Comments tell people what the code does, not what you want to do**

- Refers to things in paper, e.g. `## effect size, Equation 14 in Cohen (19XX)`
- Separates chunks of code, e.g. `## analyse reaction times` or `## preprocessing`
- Explains purpose of code, e.g. `## run costum function for t-test on each ppt`
- They live in your R script.
- Don't overdo it. Bad example is

```{r}
# Set x to 5
x <- 5
```


---

class: middle left inverse

# Roxygen2

.pull-left[

Cons:

- The **source code** (your R script) is where the code and only the code lives.
- Only have rudimentary support for documentations. 

Pros:

- …but still better than nothing.
- Good for short functions you came up in your analysis.

]

.pull-right[

[roxygen2](https://roxygen2.r-lib.org/) will turn your in-line comments beginning with `#'` to R documentation.

```{r}
#' Add together two numbers
#'
#' @param x A number
#' @param y A number
#' @return The sum of \code{x} and \code{y}
#' @examples
#' add(1, 1)
#' add(10, 1)
add <- function(x, y) {
  x + y
}
```

Still, don't write an essay!

]


.footnote[
.bold[*] Check out this article for some more cons and pros on the [roxygen way](https://www.r-bloggers.com/2018/02/documenting-r-packages-roxygen2-vs-direct-rd-input/)
]

???
roxygen2 is the developers' way to not do real documentation but only use comments

---

class: middle left inverse

# Code Styling

- Consistent spacing.
- Line length is is below 100.
- Consistent naming with `snake_case`. 

Have a look at the following resources:

- [Google's R Style Guide](https://google.github.io/styleguide/Rguide.html)
- [Tidyverse Style Guide](https://style.tidyverse.org/)

.footnote[
.bold[*] - Fun article about looking at code: [Code Review Withouth Your Glasses](https://robertheaton.com/2014/06/20/code-review-without-your-eyes/)

]

---

class: middle left inverse

# Functional Programming

Get all even numbers up to 200000 using S-style vector allocation:

.pull-left[

**Imperative-style programming**

```{r, eval = FALSE}
x <- c()
for(i in 1:200000){
    if(i %% 2 == 0){
        x <- c(x, i)
    }
}
```

]

.pull-right[

**FP-style vectorised operation**

```{r, eval = FALSE}
a <- 1:200000
x <- a[a %% 2 == 0\]
```

**Slow tidyverse method**

```{r, eval = FALSE}
## or slow tidyverse
x <- tibble(i = 1:200000) %>%
  filter(i %% 2 == 0)

```

]

.footnote[
.bold[*]
Check out the book [Functioncal Programming in R](http://adv-r.had.co.nz/Functional-programming.html).

.bold[*]
More examples can be found in [Advanced R](http://adv-r.had.co.nz/Functional-programming.html)

]

---

class: middle left inverse

# Programming with Functions

.pull-left[

```{r, eval = FALSE}
library(BayesFactor)

## subset data
test_A <- dta %>%
  filter(stimuli == "A") %>%
  summarise(mean(response))
## test against chance
ttestBF(formula = test_A$prob, mu = 0.50)

## subset data
test_B <- dta %>%
  filter(stimuli == "B") %>%
  summarise(mean(response))
## test against chance
ttestBF(formula = test_B$prob, mu = 0.50)

## subset data
test_C <- dta %>%
  filter(stimuli == "C") %>%
  summarise(mean(response))
## run test against chance
ttestBF(formula = test_C$prob, mu = 0.50)
```

This goes on for 4 more times.

]

.pull-right[

```{r, eval = FALSE}
library(BayesFactor)

costum_BF <- function(data, stimulus, null) {
  ## run test against chance
  outcome <- ttestBF(x = data$prob, mu = null)
  out <- data.frame(outcome)$bf
  ## tidy up output
  out_data <- cbind(stimuli = unique(data$abstim), bf = out, null = null)
  out_data <- data.frame(out_data) ## group_modify needs a data.frame
  return(out_data)
}


## apply test to each stimuli
tdta %>%
  group_by(abstim) %>%
  group_modify(~ costum_BF(data = .x, stimulus = Groups, null = 0.50))

```

It is easy to read 👓 and easier to debug ⚠️  or hack 👺.

]

.footnote[.red.bold[*] Apparently, the 5th circle of R Hell is reserved for people who do not write functions, check out [R Inferno](https://www.burns-stat.com/pages/Tutor/R_inferno.pdf)]

---

class: middle left inverse

# Exercise 1

1. Grab your *incredibly old* R script from a year ago, maybe even try to `source("my_file.R")` to see if you can still run it without any human intervention.

2. `linter` your code. Simply take the code below and run it:

```{r, eval = FALSE, error = FALSE}
## particularly clumsy loading of package
if (!any("lintr" %in% installed.packages()[, 'Package'])) {
  install.packages("lintr")
} else {
  library("lintr")
}

## linter your code
lintr::lint("path/to/your_file.R")
```

---

class: middle left inverse

# Exercise 2

<!-- TODO: add link to bad script -->

Here, you will look through a script and do three things:

1. **Comment** and **Prettify** code as we discussed it before
2. **Write** a function to get rid of redundant code
3. **Document** said function with `roxygen2`

---

class: middle left inverse

# Resources

Here are a list of resources to help you become your better self (*unashamed new-wave bookshop sales pitch*):

1. The Turing Way
2. renv
3. Functional Programming in R
4. Google's R Style Guide

---

class: middle left inverse

# The End

`r emo::ji("email")` lenard.dome@plymouth.ac.uk

slides are availabe at [lenarddome.github.io/cleanRcode](link)

## Potential Future Extensions

- `r fa("code-branch")` Version Control
- `r emo::ji("whale")` R Environments and maybe Dockers
- `r emo::ji("check")` Independent Tests.
